# Intro to Data Science - HW 6
##### Copyright Jeffrey Stanton, Jeffrey Saltz, and Jasmina Tacheva


```{r}
# Enter your name here: 
```


### Attribution statement: (choose only one and delete the rest)


```{r}
# 1. I did this homework by myself, with help from the book and the professor.
# 2. I did this homework with help from the book and the professor and these Internet sources:
# 3. I did this homework with help from <Name of another student> but did not cut and paste any code.
```

Last assignment we explored **data visualization** in R using the **ggplot2** package. This homework continues to use ggplot, but this time, with maps.  In addition, we will merge datasets using the built-in **merge( )** function, which provides a similar capability to a **JOIN in SQL** (don't worry if you do not know SQL). Many analytical strategies require joining data from different sources based on a **“key”** – a field that two datasets have in common. 

## Step 1: Load the population data

A.	Read the following JSON file, https://intro-datascience.s3.us-east-2.amazonaws.com/cities.json and store it in a variable called **pop**.

Examine the resulting pop dataframe and add comments explaining what each column contains.



```{r}
library(jsonlite)

# Read the JSON file
pop <- fromJSON("https://intro-datascience.s3.us-east-2.amazonaws.com/cities.json")

# Examine the structure of the dataframe
str(pop)
```

B.	Calculate the **average population** in the dataframe. Why is using mean() directly not working? Find a way to correct the data type of this variable so you can calculate the average (and then calculate the average)

Hint: use **str(pop)** or **glimpse(pop)** to help understand the dataframe


```{r}

pop$population <- as.numeric(pop$population)

average_population <- mean(pop$population, na.rm = TRUE)
average_population
```

C.	What is the population of the smallest city in the dataframe? Which state is it in?


```{r}
min_population <- min(pop$population, na.rm = TRUE)
smallest_city <- pop[pop$population == min_population, ]
smallest_city
```

## Step 2: Merge the population data with the state name data

D)	Read in the state name .csv file from the URL below into a dataframe named **abbr** (for "abbreviation") – make sure to use the read_csv() function from the tidyverse package: <br>
https://intro-datascience.s3.us-east-2.amazonaws.com/statesInfo.csv



```{r}
install.packages("tidyverse")
library(tidyverse)

# Read the CSV file into a dataframe
abbr <- read_csv("https://intro-datascience.s3.us-east-2.amazonaws.com/statesInfo.csv")
```

E)	To successfully merge the dataframe **pop** with the **abbr** dataframe, we need to identify a **column they have in common** which will serve as the **“key”** to merge on. One column both dataframes have is the **state column**. The only problem is the slight column name discrepancy – in **pop**, the column is called **“state”** and in **abbr** – **“State.”** These names need to be reconciled for the merge() function to work. Find a way to rename **abbr’s “State”** to **match** the **state column in pop**.    


```{r}
abbr <- rename(abbr, state = State)
```

F)	Merge the two dataframes (using the **‘state’ column** from both dataframes), storing the resulting dataframe in **dfNew**.


```{r}
dfNew <- merge(pop, abbr, by = "state")
```

G)	Review the structure of **dfNew** and explain the columns (aka attributes) in that dataframe.
The dfNew dataframe now has:
state: State abbreviations.
city: City names.
population: Population of each city.
Additional columns from abbr that provide more detailed information about each state.

```{r}
str(dfNew)
```

## Step 3: Visualize the data

H)	Plot points (on top of a map of the US) for **each city**. Have the **color** represent the **population**.


```{r}
library(ggplot2)
install.packages("maps")
library(maps)

# Base map of the US
us_map <- map_data("state")

# Plot points for each city
ggplot() +
  geom_polygon(data = us_map, aes(x = long, y = lat, group = group), fill = "white", color = "black") +
  geom_point(data = dfNew, aes(x = longitude, y = latitude, color = population), size = 2) +
  scale_color_viridis_c() +
  labs(title = "US Cities by Population") +
  theme_minimal()
```

I)	Add a block comment that criticizes the resulting map. It’s not very good.


```{r}
#This map isn’t very informative for identifying city populations because the #color scale is too broad, making it difficult to see differences
#the city points may be crowded in densely populated areas


```

## Step 4: Group by State


J)	Use group_by and summarise to make a dataframe of state-by-state population. Store the result in **dfSimple**.


```{r}
dfSimple <- dfNew %>%
  group_by(state) %>%
  summarise(total_population = sum(population, na.rm = TRUE))

dfSimple
```

K)	Name the most and least populous states in **dfSimple** and show the code you used to determine them.


```{r}
most_populous <- dfSimple[which.max(dfSimple$total_population), ]
least_populous <- dfSimple[which.min(dfSimple$total_population), ]

most_populous
least_populous
```

## Step 5: Create a map of the U.S.,  with the color of the state representing the state population

L) Make sure to expand the limits correctly and that you have used **coord_map** appropriately.



```{r}
library(ggplot2)
library(dplyr)
library(maps)

state_map <- map_data("state") %>%
  rename(state = region) %>%
  left_join(dfSimple, by = "state")

ggplot(state_map, aes(x = long, y = lat, group = group, fill = total_population)) +
  geom_polygon(color = "yellow", size = 0.3) +
  coord_map() +
  scale_fill_viridis_c(option = "plasma", direction = -1, 
                       name = "Population", 
                       labels = scales::comma) +
  labs(
    title = "Population Distribution by State in the U.S.",
    subtitle = "States colored by total population, highlighting disparities",
    caption = "Data Source: U.S. Cities and States"
  ) +
  theme_minimal() +
  theme(
    legend.position = "right",
    plot.title = element_text(size = 16, face = "bold"),
    plot.subtitle = element_text(size = 12),
    plot.caption = element_text(size = 10)
  )
```
